---
title: "Forecasting and Time Series Analysis Using R: Exploratory Analysis"
author: "Illarion  Jabine"
date: "6/4/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### 1. Required packages:

* forecast: Methods and tools for displaying and analysing univariate time series forecasts including exponential smoothing via state space models and automatic ARIMA modelling.
* xts: Extensible time series class that provides uniform handling of many R time series classes by extending zoo.
* astsa: Applied Statistical Time Series Analysis
* corrplot: A graphical display of a correlation matrix or general matrix. It also contains some algorithms to do matrix reordering.
* PerformanceAnalytics: Econometric Tools for Performance and Risk Analysis
* TTR: Technical Trading Rules
* GGally: Extension to 'ggplot2'


### 2. Key terms
 * Forecast 
 * Time Series
 * Forecasting Steps

### 3. Useful Links & books
 $ <https://cran.r-project.org/web/views/TimeSeries.html>: CRAN Task View: Time Series Analysis
 $ <https://otexts.com/fpp2/>: Forecasting: Principles and Practice by Rob J Hyndman and George Athanasopoulos
 $ <https://www.stat.pitt.edu/stoffer/tsda/>
 $ <https://www.crcpress.com/Time-Series-A-Data-Analysis-Approach-Using-R/Shumway-Stoffer/p/book/9780367221096>: Time Series: A Data Analysis Approach Using R by Robert Shumway, David Stoffer
 $ <https://github.com/joshuaulrich/xts>: Github of the xts package creator. 
 
### 4. Introduction

 The first thing in forecasting project is to set up our time series in R by loading the data and creating a xts object. After that explaratory analysis can be started. Visual analytics enable to spot patterns, unusual observations, changes over time, and relationships between variables. Then all these observations must be incorporated into the forecasting methods.


### 5. Load the libraries
Let's first load the libraries.
```{r loading packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(xts)
library(corrplot)
library(forecast)
library(GGally)

```

### 6. Loading and checking the data

 6.1 Load and check the datasets.
  Dataset1: transactions.csv: detailed daily sales transactions including customer   information, product, amout and sales chanel.
  Dataset2: articles.csv: product hierarchy.
 
```{r load the data and pre-process them}
# Loading data from Rds file
transactions <- read_csv("transactions.csv")
articles <- read_csv("articles.csv")

# Checking if there are any NAs:
apply(transactions,2,anyNA)
apply(articles,2,anyNA)

# There are 6 NAs in Client City transactions. 

sum(is.na(transactions$client_city))

# I need to add the product hierarchy to transactions dataset from articles dataset:

transactions <- inner_join(transactions,articles,by = "article")
# Majority of transactions belong to TRAIN category:
barplot(table(transactions$article_category))
# I will need the product hierarchy later on for botton-up and botton-down forecasting.

```

If your data is in the text file the xts object can be created with the following command:
data <- read.zoo("file.csv", header = TRUE, sep = ",",FUN = as.Date)
data <- as.xts(data)
Note: the first column inn file.csv must contain the date.
As I have already loaded the data into transactions data frame, I need to use a construct xts() to convert it into the time series xts object.
Important, as xts is a matrix it must contain numeric values!
As transactions data frame contains detailed daily sales transactions (with one particular day having many transactions) I need to aggregate them and calculate the daily turnover. Remember data points in a time series must be unique, so if daily frequency then only one point per day, etc. If I had an hourly or by minute points then we can have several different points per day.
I will calculate several aggregates, widen them, i.e. increasing the number of columns and decreasing the number of rows and then convert them into xts time series.

```{r, aggregate and widening the data frames}

# 1. Let's create aggregated data frames
aggregate_by_date <- transactions %>% group_by(date) %>% summarise(amount = sum(amount))
aggregate_by_client_city <- transactions %>% group_by(date,client_city) %>% summarise(amount = sum(amount))
aggregate_by_client_country <- transactions %>% group_by(date,client_country) %>% summarise(amount = sum(amount))
aggregate_by_article <- transactions %>% group_by(date,article) %>% summarise(amount = sum(amount))
aggregate_by_sales_chanel <- transactions %>% group_by(date,sales_chanel) %>% summarise(amount = sum(amount))
aggregate_by_article_category <- transactions %>% group_by(date,article_category) %>% summarise(amount = sum(amount))

#2. Now all aggregated data frames except the first one I need to widen, i.e. factor variables like client country or article stored in each individual row, will go to columns now.
# To widen I will use pivot_wider() (replaced spread function) from tidyr package.
# The result will be an individual time serie for each categorical variable:

wide_by_client_city <- pivot_wider(aggregate_by_client_city,names_from = client_city,values_from = amount,values_fill = list(amount = 0))

wide_by_client_country <- pivot_wider(aggregate_by_client_country,names_from = client_country,values_from = amount,values_fill = list(amount = 0))

wide_by_article <- pivot_wider(aggregate_by_article,names_from = article,values_from = amount,values_fill = list(amount = 0))

wide_by_sales_chanel <- pivot_wider(aggregate_by_sales_chanel,names_from = sales_chanel,values_from = amount,values_fill = list(amount = 0))

wide_by_article_category <- pivot_wider(aggregate_by_article_category,names_from = article_category,values_from = amount,values_fill = list(amount = 0))

wide_by_client_country <- pivot_wider(aggregate_by_client_country,names_from = client_country,values_from = amount,values_fill = list(amount = 0))

```
 
 6.2 Creating time series xts object
 
 An xts object extends the S3 class zoo from the package of the same name. It is a special class of object to handle time series.
xts object contains the time index and the matrix with actual data. So, it is a matrix with associated times for each observation.
The index is a time object: Date,POSIX Time,timeDate, chron

```{r creating xts time series objects}
ts_global <- xts(aggregate_by_date[-1], order.by = as.Date(aggregate_by_date$date))
ts_client_city <- xts(wide_by_client_city[-1], order.by = as.Date(wide_by_client_city$date))
ts_client_country <- xts(wide_by_client_country[-1], order.by = as.Date(wide_by_client_country$date))
ts_article <- xts(wide_by_article[-1], order.by = as.Date(wide_by_article$date))
ts_sales_chanel <- xts(wide_by_sales_chanel[-1], order.by = as.Date(wide_by_sales_chanel$date))
ts_article_category <- xts(wide_by_article_category[-1], order.by = as.Date(wide_by_article_category$date))
ts_client_country <- xts(wide_by_client_country[-1], order.by = as.Date(wide_by_client_country$date))

```

 6.3 Working with time series
 
As time series in its core is a matrix, we can apply various R commands and math operations: (+, -, *, /) 
We can see index and data elements of the time series as well as summary, head and tails, etc...

```{r}
# index (time) element index() is used to extract the index a.k.a. times
index(ts_global)

# coredata() is used to extract the data component
coredata(ts_global)

# head and tail:
head(ts_article_category)
tail(ts_article_category)

# summary:
summary(ts_article_category)

# Standard R functions also works with time series:
colSums(ts_article_category)
colMeans(ts_article_category)
apply(ts_article_category,2,hist)

mean(ts_global)
sd(ts_global)
var(ts_global)
hist(ts_global)
```
xts object supports comprehensive subsetting. Subsetting via "[" methods offers the ability to specify dates by range

```{r subsetting time series}
# show me only january and february 2016 and "WEB" and "WER" sales channels:

ts_sales_chanel["2016-01/2016-02",c("WEB","WER")]

```
xts package also offers other methods to manipulate time series including: merge, cbind, rbind, etc.

### 7. Time series plots

forecast package offers some time series visualisation functions, but unfortunately they only work with the standard R time series object of class "ts". So I have to convert my xts object into ts one. I will just do it for two time series.
It is also important to note here that many standard R functions do not work with xts object, so you have to convert it first to ts object.
Another important factor when creating a time series object is frequency.
Frequency is determined by domain knowledge and here I assume it to be monthly, frequency = 12.
The “frequency” is the number of observations before the seasonal pattern repeats.1 When using the ts() function in R, the following choices should be used:

Data/frequency
Annual/1
Quarterly/4
Monthly/12
Weekly/52

```{r converting to ts object}
ts_class_global <- ts(aggregate_by_date[-1],frequency = 12)
ts_class_article_category <- ts(wide_by_article_category[-1],frequency = 12)
ts_class_sales_chanel <- ts(wide_by_sales_chanel[-1],frequency = 12)
# let's check the frequency
frequency(ts_class_global)

```

We can decompose time series into trend, seasonal and random components (again decomposing only works with time series of class ts):

```{r decomposing time series}
plot(decompose(ts_class_global))
```

7.1. Univariate time series analysis

1. Standard statistical plots like histograms, box plots and qq plots can be produced for time series:

```{r time series histograms and box plots}

hist(ts_global, main = "Sales Histogram", probability = TRUE)

# Add a density line to the histogram
lines(density(ts_global),col = "red", lwd = 2)

# boxplot() only works with ts class object:
boxplot(ts_class_global, horizontal = TRUE, main = "Sales 2015 - 2019")
abline(v = median(ts_class_global), col = "red")

# QQ plot
qqnorm(ts_class_global)
qqline(ts_class_global, col = "red")
```


Time plots are great for visualising individual time series or exploring relationships between time series.

2. A time plot is where observations are plotted against the time of observation, with consecutive observations joined by straight lines. To produce a simple time plot use plot() or plot.xts():
```{r simple time plot}
plot(ts_sales_chanel["2015","WEB"], main = "Sales", xlab = "Year", ylab = "Euro", col = "blue", lwd = 2)

# We can add another time series to the same plot:
lines(ts_sales_chanel["2015","WER"], col = "red", type = "h")

# abline() does not work with xts time series, but work with ts class:

plot(ts_class_global)
abline(h = mean(ts_class_global), col = "red", lty = "dashed", lwd = 2)

```

autoplot() automatically produces an appropriate plot of whatever you pass to it in the first argument:

```{r}
autoplot(ts_global) +
  ggtitle("Sales by Year") +
  xlab("Year") +
  ylab("Thousands")

autoplot(ts_client_country["2018",c("BE","RU")]) +
  ggtitle("Sales by Year") +
  xlab("Year") +
  ylab("Thousands")

```


3. A seasonal plot ggseasonplot() (only works with ts object) is similar to a time plot except that the data are plotted against the individual “seasons” in which the data were observed.

```{r}
ggseasonplot(ts_class_global, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("EUR") +
  ggtitle("Seasonal plot: Sales")

# polar plot
ggseasonplot(ts_class_global, polar = TRUE) +
  ylab("EUR") +
  ggtitle("Seasonal plot: Sales")

```

4. Seasonal subseries plots (only works with ts object) - seasonal patterns is where the data for each season are collected together in separate mini time plots.

```{r}
ggsubseriesplot(ts_class_global) +
  ylab("Eur") +
  ggtitle("Seasonal subseries plot: sales")
```

7.2. Multivariate time series analysis

Time series with higher dimensions also known as multivariate time series bring additional possibilities to identify how they interact and explore relationships between time series. For example, how changes in marketing budget bring changes in sales volume in one or several sales channels. If we have small number of time series we still can display them on one plot.
However, if you have hundreds of dimensions it can quickly become impossible to plot them all on one single time series plot. To see the interaction between time series other techniques are used such as variance and covariance matrix, correlation plots or heatmaps, etc.
We can visualize as many time series on one plot as we want. In my examples I am using standard R plotting functionality, using ggplot2 you can create more complex graphics:

```{r 2 time series on one plot}
plot(ts_sales_chanel$WEB["2017"], main = "WEB and WER 2017 sales", col = "red")
lines(ts_sales_chanel$WER["2017"], col = "blue")
axis(side = 4, at = pretty(ts_sales_chanel$WER["2017"]))
legend(x = "right",
       legend = c("WEB", "WER"),
       col = c("red", "blue"),
       lty = c(1, 1))
```

Or using autoplot() wrapper function for ggplot2:

```{r autoplot}

# On two separate plots
autoplot(ts_sales_chanel["2017",c("WEB","WER")]) +
  xlab("Year: 2017") + ylab("") +
  ggtitle("WEB and WER 2017 Sales")

# On one plot
autoplot(ts_sales_chanel["2017",c("WEB","WER")], facets = FALSE) +
  xlab("Year: 2017") + ylab("") +
  ggtitle("WEB and WER 2017 Sales")

```

As we now analyzing multivariate time series we can plot one series against the other using scatterplots. Do not forget that a time series object is just enhanced data matrix, aware of time through tme index attached to it.
To make it more interesting I will merge two multidimensial time series: ts_article_category and ts_sales_chanel using merge() function and produce a scatterplot, which can help us to visualise the relationship between the variables:

```{r}
ts_merged_article_sales_chanel <- merge(ts_article_category,ts_sales_chanel,join = 'inner')

vis_plot <- qplot(WEB,TRAIN,data = ts_merged_article_sales_chanel) +
  ylab("TRAIN") +
  xlab("WEB")

# We can add regression line as well
vis_plot + geom_smooth(method = lm, se = FALSE)
```

When there are several predictors, it is good to plot them against each other by using scatterplot matrices. Standard pairs() function can be used or more comprehensive ggpairs() from GGally package. The scatterplot matrix enables a quick view of the relationships between all pairs of variables.

```{r}
# A convenient way to compare pairs of time seriesa. Pairs is a  facetted scatterplot of all pairs shown at once in a matrix format.
pairs(ts_class_sales_chanel[,c("WEB","WER","GSB")],lower.panel = NULL,main = "Sales Channel Correlation Scatterplots")

ggpairs(as.data.frame(ts_class_sales_chanel[,c("WEB","WER","GSB")]))

# We can calculate a correlation matrix
cor_matrix <- cor(ts_class_sales_chanel)

# And plot this matrix using corrplot() from corrplot package. It allows to represent the correlations with numbers, symbols, colors etc.
corrplot(cor_matrix,method = "number",
type = "upper")

# The same corrplot but in a form of color heatmap 
corrplot(cor_matrix,method = "color",
type = "upper")

# corrplot() offers some intellegent way to cluster matrix cells using hierarchical clustering:
corrplot(cor_matrix,method = "color",
type = "upper", order = "hclust")

```

7.3 Autocorrelation

Autocorrelation (or self-correlation) is one of the fundamental ideas in time series analysis. Autocorrelation indicates the relationship between one point in the time series and points that come before it. Autocorrelation measures the linear relationship between lagged values of a time series.
The autocorrelogram is a chart which indicates the correlation between points separated by various time lags. 
By default, it displays the first 30 lags: the correlation between points n and n - 1 (lag 1), n and n - 2 (lag 2), n and n - 3 (lag 3) and so on up to 30. 
The autocorrelogram shows how any point in the time series is related to its past and the significance of this relationship.The significance levels are given by 2 horizontal lines above and below 0.
To produce autocorrelation chart run acf() function (or ggAcf() "nicer" ggplot2 wrapper function from forcast package), or you can calculate autocorrelation manually using standard R function cor():

```{r}
# lag 1 autocorrelation using cor():
cor(ts_global[-length(ts_global)],ts_global[-1])

# lag 2:
cor(ts_global[-c(length(ts_global),length(ts_global)-1)],ts_global[-c(1,2)])

# Or using acf() function, if you do not want to see teh plot set plot = FALSE:
acf(ts_global,plot = FALSE)

# To produce the autocorrelation chart:
acf(ts_global)

# or ggAcf()
ggAcf(ts_global, lag.max = 50)


```
By looking at the autocorrelation plot you can tell a lot about properties and patterns (trend and seasonality) existing in the time series.
Note that acf at lag-0 is always 1.
In this example we have peaks at r7, r14, r21. Can we conclude that there is a seasonal pattern in the data repeating every 7 days?
The dashed blue lines indicate whether the correlations are significantly different from 0, i.e. no autocorrelation at that lag.

Hyndman et al point out the following properties of ACF:
 1. "When data have a trend, the autocorrelations for small lags tend to be large and positive because observations nearby in time are also nearby in size. So the ACF of trended time series tend to have positive values that slowly decrease as the lags increase."
 2. "When data are seasonal, the autocorrelations will be larger for the seasonal lags (at multiples of the seasonal frequency) than for other lags."
 3. "When data are both trended and seasonal, you see a combination of these effects."

### 8. Additional xts features

xts provides extra features that allow to find repeating itervals, identify frequency and coverage in time, etc.

 8.1. Determining periodicity
Periodicity defines regularity with what your data repeat. periodicity() function estimates the periodicity of a time-series-like object by calculating the median time between observations in days. This is the first step to understanding your data before going into further analysis. periodicity() also returns the range of your time series index. 
```{r periodicity}
periodicity(ts_article)
```

You can convert periodicity using to.period() xts functions:
to.minutes(x,k,name,...)
to.minutes3(x,name,...)
to.minutes5(x,name,...)
to.minutes10(x,name,...)
to.minutes15(x,name,...)
to.minutes30(x,name,...)
to.hourly(x,name,...)
to.daily(x,drop.time=TRUE,name,...)

 8.2. Finding the number of periods in TS
To know how many discrete irregular periods your time series data covers, use the function ndays() and its shortcut functions nmonths(), nquarters(), etc.
```{r number of periods}
# Count the months
nmonths(ts_global)

# Count the quarters
nquarters(ts_global)

# Count the years
nyears(ts_global)

```
