---
title: "Manipulating Time Series: Math, Functions, Merging etc..."
author: "Illarion  Jabine"
date: "9/4/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### 1. Required packages:

* xts: Extensible time series class that provides uniform handling of many R time series classes by extending zoo.


### 2. Key terms
 * TS subsetting 
 * Time based queries
 * Merging TS
 * Rolling functions
 * Apply functions by time

### 3. Useful Links & books
 $ <https://cran.r-project.org/web/views/TimeSeries.html>: CRAN Task View: Time Series Analysis
 $ <https://www.iso.org/standard/40874.html>
 $ <https://github.com/joshuaulrich/xts>: Github of the xts package creator. 
 
### 4. Introduction

In this manual I will cover different methods and approaches of how to work with xts time series objects. In particular math operations, time based queries, time series functions, merging, subsetting.

### 5. Load the libraries
Let's first load the libraries.
```{r loading packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(xts)
library(lubridate)

```

### 6. Loading and checking the data

I will use the time series in xts format that have been created at the previous step (see my manual "2.Forecasting-Exploratory-Analysis.Rmd").
For the conviniens I have saved all the time series objects as a RData file.
Note: RDS files can store a single R object, and RData files can store multiple R objects.
To save all time series objects use this command: save(ts_global,ts_article,ts_article_category,ts_client_city,ts_client_country,ts_sales_chanel, file = "time_series.RData"))

 
```{r load the data from RData file}
# Loading data from Rds file
load("time_series.RData")

plot(ts_article["2017",])

```


### 7. Time based queries

xts supports R time based queries and ISO 8601:2004 standard. This standard defines representation of dates and times.
xts object supports both standard R matrix subsetting and using character strings that adhere to the ISO-8601 standard:
 * One and two sided intervals "2004" & "2001/2015"
 * Truncated representation "201402/03"
 * Time support "2014-02-22 08:30:00"
 * Repeating intervals "T08:00/T09:00"
What is also important is that all subsets preserve matrix (drop = FALSE) and order is also preserved.
 1. Time based queries
```{r subsetting xts object}

# April 2017
plot(ts_global["2017"])

# January 2018 to March:
plot(ts_article["201801/03",c("SNCF","THALYS")])
# or the same result: plot(ts_article["2018-01/2018-03",c("SNCF","THALYS")])


# Select January 1, 2017 to February 15, 2017
plot(ts_global["2017/2017-02-15"])

# If time series has been captured per minute periodicity xts supports data time: YYYYMMDDTHHMM format

# Let's genererate time index
idx <- ymd_hms(c("2018-02-15 20:30:30","2018-02-15 20:31:00","2018-02-15 20:31:30","2018-02-15 20:32:00"))

# Now generate the actual data
x <- rnorm(4)

# Construct xts object
ts_x <- xts(x, order.by = idx)

# Show observations done on 2018-02-15 at 20:31
ts_x["2018-02-15T2031"]

#Repeating intraday intervals

ts_x["T20:31/T20:31"]

# Extract all data between 7AM and 11AM
ts_x["T07:00/T11:00"]

```

 2. Information from a time series can be extracted using alternative extraction techniques.

```{r}
#Integer indexing
ts_article[1:10,c("SNCF")]

ts_global[c(2,4,6)]

# Logical vectors using index() method
ts_global[index(ts_global) > "2019-01-01" & index(ts_global) < "2019-01-05"]

ts_article[index(ts_global) > "2019-01-01","SNCF"]

```

 3. Update and replace elements in xts object
It is easy to update values in xts object.
To change individual observations located sporadically throughout a time series, integers or logical vectors can be used.

```{r replacing values in xts objects}

# Let's create a date index
date_idx <- seq(as.Date("2020-03-01"), length = 50, by = "days")

# and create a xts object
ts_obj <- xts(1:50, order.by = date_idx)

dates <- c("2020-04-15","2020-04-16")
ts_obj[dates] <- NA

ts_obj["2020-03-01"] <- 1000
```


 4. Finding the first or last period of time
Use standard head() and tail() or 2 similar xts functions with respect to time: first() and last().
If you need to locate data by relative time, but instead of using an absolute offset, you use a relative time position. For example, the last 2 weeks of a series, or the first day of current month. 
xts object is aware of time and supports relative time position by means of first() and last() methods.
These two methods support the following character contexts:
seconds,
minutes,
hours,
days,
weeks,
months,
quarters,
years
Let's extract relative observations from ts_global time series:

```{r relative time extracts}

# first 3 weeks
first(ts_global, "3 weeks")

# last two months of 2018
plot(last(ts_global["2018"],"2 month"))

# last quarter of 2017
last_quarter_2017 <- last(ts_global["2017"],"1 quarters")

# Extract all but the first 4 days of last quarter of 2017

first(last_quarter_2017, "-4 days")

```

You can build complex chains and extract a very specific subset of observations by linking together multiple calls to first() and last().

```{r}
# Last 5 days of first week
last(first(ts_global, '1 week'), '5 days') 

# First three days of the second week
first(last(first(ts_global, "2 weeks"), "1 week"), "3 days")

```


### 8. Time series arithmetic

xts object is a matrix, so most rules applicable to matrix operations are also applicable to xts object.
Naturally, you can always add a column to your time series object, or perform other math operations.
It is important to note that math operations are on the intersection of times, i.e. only these intersections will be used

```{r}
# adding a new column
ts_article_category$New_Column <- ts_article_category$TRAIN * 2

# Deleting the new column:
ts_article_category$New_Column <- NULL

ts_math <- first(ts_article["2018",c("SNCF","THALYS","EUROSTAR")],"1 month")

# +,-,*,/

# add 1000 element wise
add <- ts_math + 1000

# multiply all elements of the matrix by 10 
multiply_10 <- ts_math * 10

# divide one time series by another (the two TSs have the same time index)
multiply_10/ts_math

# If I have two time series but with slightly different time index then only intersections will be used:

x <- xts(c(1,1,1),order.by = as.Date(c("2020-01-01","2020-01-02","2020-01-03")))
y <- xts(c(2,2,2),order.by = as.Date(c("2020-01-01","2020-01-02","2020-01-04")))

# Intersection of dates when x + y

x + y
```

### 9. Merging time series

