---
title: "Manipulating Time Series: Math, Functions, Merging etc..."
author: "Illarion  Jabine"
date: "11/4/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### 1. Required packages:

* xts: Extensible time series class that provides uniform handling of many R time series classes by extending zoo.


### 2. Key terms
 * Lagging
 * Difference
 * Rolling functions
 * Apply a function by time period(s)

### 3. Useful Links & books
 $ <https://cran.r-project.org/web/views/TimeSeries.html>: CRAN Task View: Time Series Analysis
 $ <https://www.iso.org/standard/40874.html>
 $ <https://github.com/joshuaulrich/xts>: Github of the xts package creator. 
 
### 4. Introduction

In this manual I will cover different methods and approaches of how to apply different functions by time to xts time series objects. In particular lag, diff, rolling, aggregate and apply functions.
Practical examples can be: average/cumulative sales per period (week, month, etc)

### 5. Load the libraries
Let's first load the libraries.
```{r loading packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(xts)
library(lubridate)

```

### 6. Loading and checking the data

I will use the time series in xts format that have been created at the previous step (see my manual "2.Forecasting-Exploratory-Analysis.Rmd").
For the conviniens I have saved all the time series objects as a RData file.
Note: RDS files can store a single R object, and RData files can store multiple R objects.
To save all time series objects use this command: save(ts_global,ts_article,ts_article_category,ts_client_city,ts_client_country,ts_sales_chanel, file = "time_series.RData"))

 
```{r load the data from RData file}
# Loading data from Rds file
load("time_series.RData")

plot(ts_article["2017",])

```

### 7. Leading/lagging time series and difference

 7.1. To lag (aka backshift operation, Lk notation) a series is another common modification for time series. Lagging lets you see observations like yesterday's value in the context of today.
Note: lag() function from base R works slightly differently, i.e. zoo uses a convention for the sign of k in which negative values indicate lags and positive values indicate leads. In zoo lag(x, k = 1) will shift future values one step back in time.
xts implements the exact opposite, i.e. for a positive k, the series will shift the last value in time one period forward.
Note: To be 100% sure use lag.xts() function.

```{r}

x <- xts(1:10,Sys.Date() + 1:10)
# Create a leading TS
lead_x <- lag.xts(x, k = -1)

# Create a lagging TS
lag_x  <- lag.xts(x, k = 1)

# Merge three TSs together
cbind(lead_x,x,lag_x)

```

 7.2. Time series difference

 If a time series is non-stationary it can be converted to stationary one by means of difference operator. 
 A single or first order difference is calculated as x(t) - x(t-k) where k is the number of lags to go back. 
Higher order differences are simply the reapplication of a difference to each prior result.
FOr example, first order difference is when x(t) - x(t-1).
diff(x, lag = 1, differences = 1, ...) functions calculates the difference.
 * the lag argument is the number of periods to use
 * the differences is an integer indicating the order of the difference, or how many times diff() is called: 
 diff(diff(x)) is the same as diff(x, differences = 2)
```{r TS difference}

# difference can be constructed with lag operator and subtraction.
# Example of first order difference:
x - lag(x)

# it gives the same results as:
diff(x)

# If we have monthly data, we can calculate the first order 12 month difference diff(ts_monthly, lag = 12, differences = 1) x(t) - x(t-12)
```


### 8. Apply a function by time period(s)

It is extremely easy to apply functions by time in xts.

 8.1. To find out the period end points, use endpoints() function. It takes a time series and returns the locations (index) of the last observations in each interval (period).
endpoints(ts, on="months", k=1)
The argument on supports a variety of periods, including "years", "quarters", "months", as well as intraday intervals such as "hours", and "minutes".
It returns an index vector starting with 0 and ending with the extent (last row) of your data.
The k argument is used to find the Kth period. 
on = "weeks", k = 2, would generate the final day of every other week in your data. Note that the last value returned will always be the length of your input data, even if it doesn't correspond to a skipped interval.
```{r period endpoints}
# year end points
index_endpoints_year <- endpoints(ts_global,on="year")
ts_global[index_endpoints_year]

# Month end points
plot(ts_global[endpoints(ts_global,on="month")], main = "Sales on month end points")

# Locate every two weeks

plot(ts_global[endpoints(ts_global,on="week",k = 2)], main = "Sales on every second week end points")
```

 8.2. period.apply(): applying function by time period

Here we will use the endpoints and give them to the base function apply(). Then we will apply this function on a window of time.
For example, I want to know monthly total sales amount or weekly mean sales, etc.
period.apply(x, INDEX, FUN, ...) is used, where INDEX will be our endpoints.
As usual you can always pass additional function arguments to function inside apply call. 
For example, if FUN = sum, you can pass to sum() its argument na.rm = TRUE.

```{r Applying function by time period}

# monthly sales
index <- endpoints(ts_global,on="month")
ts_monthly_sales <- period.apply(ts_global, INDEX = index, FUN = sum, na.rm = TRUE)

# weekly mean sales
index_weekly <- endpoints(ts_client_country,on="week")
weekly_mean <- period.apply(ts_client_country, INDEX = index_weekly, FUN = mean)

plot(weekly_mean["201801/02","RU"])
```

 8.3. split() and lapply() to apply functions on intervals

Here is another possibility to report interval figures. 
xts split() function splits data into disjoint chunks by time. It creates a list containing an element for each split. The f argument in split() is a character string describing the period to split by: months,years,etc
Then we perform some calculations on these chunks by means of lapply().
Let's use split() to split a time seriesby month and use lapply() to get the monthly sum.
Basically we can say that period.apply() == split() + lapply()
```{r}

# Split TS by month and create a list
ts_monthly_split <- split(ts_article_category, f = "month")

# Apply sum function to the list to calculate monthly sales
ts_monthly_sales <- lapply(ts_monthly_split, sum)

# Let's check if this method produces the same result as period.apply (see 8.2)?
ts_monthly_split_global <- split(ts_global, f = "month")
ts_monthly_sales_global <- lapply(ts_monthly_split_global, sum)

# The difference is 0.
sum(unlist(ts_monthly_sales_global)) - sum(ts_monthly_sales)

# Cumulative  sum by month, here I apply cumsum() function to the list:
ts_monthly_cumsum <- lapply(ts_monthly_split, cumsum)

# let's plot the first month for TRAIN product category:
plot(ts_monthly_cumsum[[1]]$TRAIN, main = "1 month TRAIN cum sum")

# Let's validate the result, must be 0:
sum(cumsum(ts_article_category["201501","TRAIN"]) - ts_monthly_cumsum[[1]]$TRAIN)
```

### 9. Rolling Functions 

Another common task while working with time series data is to apply a function on a rolling window of data.
In zoo it's implemented through rollapply() function. 
It takes a time series object x, a window size width, and a function FUN to apply to each rolling period.
Attention! a number supplied to the width argument specifies the number of observations in a window. You have to be aware of the periodicity of your time series, i.e. daily, weekly, etc. SO, if width = 5, it will take the 5-day max of a series with daily observations. If the series had monthly observations, it would take the 5-month max.
Our ts_*** time series have daily observations.
Let's calculate rolling 7 day sum of sales:

```{r 7 day rolling sum}

rolling_sum <- rollapply(ts_client_country,width = 7,FUN = sum)
head(rolling_sum[,c("BE","RU")],n = 20)

# Let's calculate the cum sum and compare with what rollapply gives us:
i <- 1
j <- 7
cumsum_RU <- vector(mode = "integer")
cumsum_BE <- vector(mode = "integer")
while (i < 15) {
  while (j < 22) {
  cumsum_RU[i] <- sum(ts_client_country[i:j,"RU"]) 
  cumsum_BE[i] <- sum(ts_client_country[i:j,"BE"])
  j <- j + 1
  break
    }
 i <- i + 1
}
res <- cbind(cumsum_BE,cumsum_RU)

# validating and the difference must be 0:
sum(res) - sum(head(rolling_sum[,c("BE","RU")],n = 20), na.rm = TRUE)


```
 
 
