---
title: 'Customer Analytics usinf R'
author: "Illarion  Jabine"
date: "10/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### 1. Required packages:

* []: 
* []:

### 2. Key terms
 * 
 * 
 * 

### 3. Useful Links
 $ Wharton Customer Analytics Initiative: <https://wcai.wharton.upenn.edu/>
 
## 4. Introduction
According to wikipedia:
Customer analytics is a process by which data from customer behavior is used to help make key business decisions via market segmentation and predictive analytics. 
This information is used by businesses for direct marketing, site selection, and customer relationship management.
Customer Analytics can provide answer to some fundamental business questions, like:
 1. How can we group/segment our customers?
 2. What are the most profitable customer segments?
 3. What customers are likely to upgrade?
 4. How can we improve our marketing campaigns?
 5. What customers are most likely to leave?
 6. What products will customers want to buy next?
 7. How to upsell/crossell, and how to bundle our products?
 8. What are customers talking about our products or services?
the list goes on....




### 5. Load the libraries
Let's first load the libraries.
```{r loading packages, message=FALSE, warning=FALSE}
library()
```

### 6. Loading and checking the data

Load and check the datasets.
Dataset1:
 * 
 * 
 
```{r load the data and pre-process them}
# Loading data from Rds file
load("data/data.Rds")

# Checking if there are any NAs:
anyNA()


```

### 7. Topic 1


### 8. Topic 2


### 9. Topic 3