---
title: 'RFM: recency, frequency, monetary marketing analysis'
author: "Illarion  Jabine"
date: "20/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### 1. Required packages:

* [rfm]: Recency, Frequency and Monetary Value Analysis

### 2. Key terms
 * RFM analysis
 * 
 * 

### 3. Useful Links
 $ <>
 
## 4. Introduction to RFM Analysis
Recency, Frequency, and Monetary Value are popular metrics used to quantify customer transaction history.
Recency is how recently the customer has purchases. Frequency is how often the customer purchases, and monetary value is the dollar value of the purchases. 
RFM analysis classifies customers into groups according to their RFM measures.
These classifications can be used to predict the customer behavior including the likelihood of responding to an offer, etc.


### 5. Load the libraries
Let's ²²first load the libraries.
```{r loading packages, message=FALSE, warning=FALSE}
library()
```

### 6. Loading and checking the data

Load and check the datasets.
Dataset1:
 * 
 * 
 
```{r load the data and pre-process them}
# Loading data from Rds file
load("data/anomaly_detection.Rds")

# Checking if there are any NAs:
anyNA(river_nitrate)
anyNA(wine)

```

### 7. Topic 1


### 8. Topic 2


### 9. Topic 3