---
title: "Six Sigma In R"
author: "Illarion  Jabine"
date: "27/01/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### 1. Required packages:

* [SixSigma]: Six Sigma Tools for Quality Control and Improvement
* [qcc]: Quality Control Charts (Shewhart quality control charts for continuous, attribute and count data.)
* [qicharts]: Quality Improvement Charts
* [qualityTools]: Statistical Methods for Quality Science <http://www.r-qualitytools.org>

### 2. Key terms
 * Six Sigma 
 * Control Charts
 * DMAIC
 * Root-Cause Diagram (Cause and Effect Diagram / Fish Bone Diagram / Ishikawa Diagram)
 * Gage R&R
 * SPC
 * Shewhart quality control charts
 * Cusum and EWMA charts
 * Operating characteristic curves
 * Process capability analysis
 * Pareto chart
 * Multivariate control charts
 * LSL – Lower Specification Limit
 * USL – Upper Specification Limit.

### 3. Useful Links
 * Six Sigma with R: <https://www.springer.com/gp/book/9781461436515>
 * Six Sigma Institute: <https://www.sixsigma-institute.org>
 * Wiki: <https://en.wikipedia.org/wiki/Six_Sigma>
 * ISO 13053-1:2011: Quantitative methods in process improvement — Six Sigma <https://www.iso.org/standard/52901.html>
 
## 4. Introduction
Six Sigma is one of the most widely used set of techniques, tools and method's aimed at 
improving business processes and optimizing operations.
Sigma is a statistical term which defines the standard deviation from from the mean of the normal distribution.
If we take 3 sigmas to left and 3 sigmas to the right from the mean (i.e. 6 sigmas in absolute value) then 99.99% of observations will be in this range.
Six Sigma is an indispensable tool for quality control and improvement initiatives.
It uses statistical methods to statistical methods to monitor and control a process.
And this is a strong point about Six Sigma.
By introducing verifiable data and statistical methods we can define, measure, analyse, improve, and control the business process. Six sigma involves systematic and data-driven approach to managing quality improvement projects.
The quality control management base their decisions on facts rather than assumptions and guesswork.
Indeed, Six Sigma clear focus on achieving measurable and quantifiable financial returns.
To improve existing business processes Six Sigma follows DMAIC and DMADV methodologies.
 * DMAIC is used for projects aimed at improving an existing business process.
 * DMADV is used for projects aimed at creating new product or process designs.
In this manual I will consider DMAIC cycle. DMAIC stands for:
 1.Define
 2.Measure
 3.Analyse
 4.Improve
 5.Control
In this manual I will only cover the topics of how R can be used in Measure, Analyse and Control phases.

### 5. Load the libraries
Let's first load the libraries.
```{r loading packages, message=FALSE, warning=FALSE}
library(SixSigma)
library(qcc)
library(qicharts)
library(qualityTools)
```

### 6. Loading and checking the data

Load and check the datasets.
Dataset1: pistonrings from qcc package
 * 
 * 
 
```{r load the data and pre-process them}
# Loading data from Rds file


# Checking if there are any NAs:


```

### 7. Define
During the definition phase we identify the problem area faced by the customer.
We identify customer requirements and set the project goals to achive. During this phase the PM prepares all the project initiation documents: problem statement, customer requiremets, project statement, business case, project charter, etc.
According to Six Sigma institute the define phase consists of the following 
SixSigma package offers the process mapping tool for define stage. Process map (<https://en.wikipedia.org/wiki/Business_process_mapping>)
Process map depicts the steps of the processes starting from a high, conceptual level and 
decomposing to lower levels. Each process consumes inputs X and produces output Y.
SixSigma package has a function ss.pMap() to draw the process map. However, the process of creating a map is quite tedious. I think it is much easier to create such a map in a graphical tool.
Here is an example of a process map from SixSigma package manual: 
```{r message=FALSE, warning=FALSE}
inputs.overall<-c("operators", "tools", "raw material", "facilities")
outputs.overall<-c("helicopter")
steps<-c("INSPECTION", "ASSEMBLY", "TEST", "LABELING")
#Inputs of process "i" are inputs of process "i+1"
input.output<-vector(mode="list",length=length(steps))
input.output[1]<-list(c("sheets", "..."))
input.output[2]<-list(c("sheets"))
input.output[3]<-list(c("helicopter"))
input.output[4]<-list(c("helicopter"))
#Parameters of each process
x.parameters<-vector(mode="list",length=length(steps))
x.parameters[1]<-list(c(list(c("width", "NC")),list(c("operator", "C")),
list(c("Measure pattern", "P")), list(c("discard", "P"))))
x.parameters[2]<-list(c(list(c("operator", "C")),list(c("cut", "P")),
list(c("fix", "P")), list(c("rotor.width", "C")),list(c("rotor.length",
"C")), list(c("paperclip", "C")), list(c("tape", "C"))))
x.parameters[3]<-list(c(list(c("operator", "C")),list(c("throw", "P")),
list(c("discard", "P")), list(c("environment", "N"))))
x.parameters[4]<-list(c(list(c("operator", "C")),list(c("label", "P"))))
x.parameters
#Features of each process
y.features<-vector(mode="list",length=length(steps))
y.features[1]<-list(c(list(c("ok", "Cr"))))
y.features[2]<-list(c(list(c("weight", "Cr"))))
y.features[3]<-list(c(list(c("time", "Cr"))))
y.features[4]<-list(c(list(c("label", "Cr"))))
y.features
ss.pMap(steps, inputs.overall, outputs.overall,
input.output, x.parameters, y.features,
sub="Paper Helicopter Project")

```


### 8. Measure
If you can not measure something how can you control it? 
Take the current process ("as is") and collect the relevant qualitative and quantative data about it.
Measuring phase can be divided into several steps:

### 8.1 Data types
Define what types of data to measure: qualitative or quantitative.
Qualitative data is subjective in nature and cannot be measured objectively. 
They can be nominal, ordinal (ranked or ordered) and binary.
Quantitative data is objective in nature and can be measured. They can be discrete or continuous

### 8.2 Type of statistics
Define what kind of statistics to gather. For quantitative data: 
    * measure of central tendency - mean, median, mode
    * measure of spread/dispertion: standard deviation, range

### 8.3 Data collection strategy.
    * Probability Sampling: 
      Simple Random Sampling
      Stratified Random Sampling
      Systematic Sampling
      Cluster Sampling
    * Non Probability Sampling: Convenience Sampling, Judgment Sampling, etc
    
### 8.4 Measurement System
The objective of the measure phase is to identify and understand the variation due to the
measurement system itself and variation arising from the actual production process.
Variation and bias are two major elements that can negatively impact the measurement process. 
Measurement System Analysis (MSA) identifies what variations come from appraiser (the operator – individual who operates the tool/instrument) and Variation from Gage due to the instrument itself. 
Variation from gage can be further classified into Accuracy, Stability, Repeatability, Reproducibility and Linearity.
The Actual Process Variation are caused by controllable factors and/or uncontrollable factors.
In general, we can calculate the observed variation as follows:
Variation (observed) = Variation (Actual) + Variation (Measurement)
Measurement system errors can be due to:
 * Accuracy – The difference between the average of observed values and the standard
 * Repeatability – Variation in measurement when a person measures the same unit repeatedly with the same measuring gage (or tool)
 * Reproducibility - Variation in measurement when two or more persons measure the same unit using the same measuring gage (or tool)
 * Stability - Variation in measurement when the same person measures the same unit using the same measuring gage (or tool) over an extended period of time.
 * Linearity – The consistency of the measurement across the entire range of the measuring gage.
Process Variation is caused by two reasons: Common Causes(noise) and Special Causes (signals).
Variation is inherent in the system and results in a stable "in control" state.
Whereas signal is an unexpected occurrenc due to unforeseen circumstances. It results in "out of control" state because the variation is not predictable or unusual.

### 8.5 Data Interpretation

Collected data can be interpreted using the following factors:
 Stability
 Normality 
 Shape
 Spread 
 Centering

Tools in R for data interpretation:
 1. Gage Repeatability and Reproducibility (Gage R&R)
 2. Run chart plots  
 3. Control charts
 4. Standard R functionality for normality check like qqplot, statistical tests, histograms, standard deviation, mean, median, etc. 
 
1. Gage Repeatability and Reproducibility
ss.rr() from SixSigma package and gageRRDesign(), gageRR() from qualityTools produce gage R&R analysis.

```{r gage R&R}
ss.rr(time1, prototype, operator, data = ss.data.rr, 
	sub = "Six Sigma Paper Helicopter Project", 
	alphaLim = 0.05,
	errorTerm = "interaction",
	lsl = 0.7,
	usl = 1.8)
```



2. Run chart
Run chart is simply a time series ordered chart of process data with mean or median shown as a horizontal line.
With the average line we can better see what is going on with the process.
Using this chart we can study trends, patterns and variation in the process data over time.
<https://cran.r-project.org/web/packages/qicharts/vignettes/runcharts.html>
```{r run chart}

# run chart using standard R functions
y <- runif(n = 20,min = 10,max = 20)
plot(y,type = "b")

# adding the horizontal line to the chart
abline(h = mean(y), v = 0, col = "blue", lty = "dashed")

# run chart using qic() from qicharts package:
qic(y)

# or tcc() from qicharts
tcc(y)
```

3. Control charts

Control charts are "enhanced" version of run charts with upper and lower control limits added.

We can build a basic control charts with standard R functions
```{r control charts: standard R}

# let's set UCL as the mean + 3 sigma
ucl <- mean(y) + 3*sd(y)

# let's define LCL as the mean - 3 sigma
lcl <- mean(y) - 3*sd(y)

plot(y,type = "b", ylim = c(4,25))

# adding the horizontal line to the chart
abline(h = mean(y), v = 0, col = "blue", lty = "dashed")

# adding the horizontal line to the chart
abline(h = round(ucl), v = 0, col = "red", lty = "dashed")
# adding the horizontal line to the chart
abline(h = lcl, v = 0, col = "red", lty = "dashed")

```

We can build advanced control charts using tcc() from qicharts, qcc() from qcc package.
There are different control charts depending on data type, on how the data were collected, etc:
Chart	  Chart description
"xbar"	means of a continuous process variable
"R"		ranges of a continuous process variable
"S"		standard deviations of a continuous variable
"xbar.one"	one-at-time data of a continuous process variable
"p"		proportion of nonconforming units
"np"	number of nonconforming units
"c"		nonconformities per unit
"u"		average nonconformities per unit
"g"		number of non-events between events
Here is a good tutorial from minitab:
<https://blog.minitab.com/blog/understanding-statistics/control-chart-tutorials-and-examples>
Also run ?qcc to read help on qcc() function.

4. Statistical tests

Test             Statistics
t.test()         mean
prop.test()      proportion
var.test()       variance
bartlett.test()  variance
chisq.test()     contingency table
poisson.test()   Poisson distribution
binom.test()     binomial distribution
shapiro.test()   normality test

### 9. Analyse
During this phase we analyse the collected data and try to find out the causes of issues.
There must be some exploitative data analysis carried out to investigate and confirm what is the causing the issues. It's an iterative approach of asking and answering questions. 

R has various statistical methods and tools to perform such analysis.
 1. Cause and effect diagram
 2. Pareto chart
 3. Bar chart
 4. Histogram
 5. Box plot
 6. Scatter plot
 7. Regression, correlation and distributions
 8. ANOVA

1. Cause and effect diagram

Also know as fishbone or Ishikawa diagrame helps visualize the reasons causing a problem.
The steps to build the diagram:
 1. Identify the main problem.
 2. Identify the major factors involved
 3. For each factor list possible causes. 
Somehow I can compare the fishbone diagram to a mind map.
ss.ceDiag() from SixSigma package and cause.and.effect() from qcc package both produce cause and effect diagram. I find it easier to use cause.and.effect() from qcc.
```{r Cause and effect diagram}
cause.and.effect(cause=list(Measurements=c("Micrometers", "Microscopes", "Inspectors"),
                            Materials=c("Alloys", "Lubricants", "Suppliers"),
                            Personnel=c("Shifts", "Supervisors", "Training", "Operators"),
                            Environment=c("Condensation", "Moisture"),
                            Methods=c("Brake", "Engager", "Angle"),
                            Machines=c("Speed", "Lathes", "Bits", "Sockets")),
                 effect="Surface Flaws")
```

2. Pareto chart

Pareto chart is a bar chart with two axes, with the left one showing the frequency count
and the right showing the cumulative percentage.
Pareto chart helps to separate the vital few causes, famous 20% of the defects cause 80 % of the problems, from the trivial many.
Once you know this vital 20% you can address them immediately.

To produce a pareto chart in R use the following functions:
 1. paretoChart() from qualityTools package
 2. paretochart() from qicharts package
 3. pareto.chart() from qcc package

```{r pareto chart using qualityTools package}
defects <- c(rep("E", 62), rep("B", 15), rep("F", 3), rep("A", 10), rep("C",20), rep("D", 10))
qualityTools::paretoChart(defects)
```


### 10. Improve

Well put in place the process improvements. 
Here is a good reading: <http://www.r-qualitytools.org/IMPROVE.html>

### 11. Control

Once the improvements have been put into productions, the performance of the new process has to be monitored and controlled proactively.
Statistical Process Control (SPC) is used to monitor and control the output parameters of the process (Y’s).
Control charts are widely used here to identify and prevent process from special causes.
