---
title: "Six Sigma In R"
author: "Illarion  Jabine"
date: "27/01/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### 1. Required packages:

* [SixSigma]: Six Sigma Tools for Quality Control and Improvement
* [qcc]: Quality Control Charts (Shewhart quality control charts for continuous, attribute and count data.)

### 2. Key terms
 * Six Sigma 
 * COntrol Chart
 * 

### 3. Useful Links
 * Six Sigma with R: <https://www.springer.com/gp/book/9781461436515>
 * Wiki: <https://en.wikipedia.org/wiki/Six_Sigma>
 * ISO 13053-1:2011: Quantitative methods in process improvement — Six Sigma <https://www.iso.org/standard/52901.html>
 
## 4. Introduction
Six Sigma is one of the most widely used set of techniques, tools and method's aimed at 
improving business processes and optimizing operations.
Sigma is a statistical term which defines the standard deviation from from the mean of the normal distribution.
If we take 3 sigmas to left and 3 sigmas to the right from the mean (i.e. 6 sigmas in absolute value) then 99.99% of observations will be in this range.
Six Sigma is an indispensable tool for quality control and improvement initiatives.
It uses statistical methods to statistical methods to monitor and control a process.
And this is a strong point about Six Sigma.
By introducing verifiable data and statistical methods we can define, measure, analyse, improve, and control the business process.
The management base their decisions on facts rather than assumptions and guesswork.
If you can not measure something how can you control it? 
Indeed, Six Sigma clear focus on achieving measurable and quantifiable financial returns.
To improve existing business processes Six Sigma follows DMAIC and DMADV methodologies.
 * DMAIC is used for projects aimed at improving an existing business process.
 * DMADV is used for projects aimed at creating new product or process designs.
In this manual I will consider DMAIC cycle. DMAIC stands for:
 1.Define
 2.Measure
 3.Analyse
 4.Improve
 5.Control

### 1. Define

During the definition phase we identify the problem area faced by the customer.
set the goals to achive

### 2. Measure


### 5. Load the libraries
Let's first load the libraries.
```{r loading packages, message=FALSE, warning=FALSE}
library(SixSigma)
library(qcc)
```

### 6. Loading and checking the data

Load and check the datasets.
Dataset1:
 * 
 * 
 
```{r load the data and pre-process them}
# Loading data from Rds file


# Checking if there are any NAs:


```

### 7. Topic 1


### 8. Topic 2


### 9. Topic 3