---
title: "Six Sigma In R"
author: "Illarion  Jabine"
date: "27/01/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### 1. Required packages:

* [SixSigma]: Six Sigma Tools for Quality Control and Improvement
* [qcc]: Quality Control Charts (Shewhart quality control charts for continuous, attribute and count data.)
* [qicharts]: Quality Improvement Charts
* [qualityTools]: Statistical Methods for Quality Science

### 2. Key terms
 * Six Sigma 
 * Control Charts
 * DMAIC
 * Root-Cause Diagram (Cause and Effect Diagram / Fish Bone Diagram / Ishikawa Diagram)
 * Gage R&R
 * SPC
 * Shewhart quality control charts
 * Cusum and EWMA charts
 * Operating characteristic curves
 * Process capability analysis
 * Pareto chart
 * Multivariate control charts
 * LSL – Lower Specification Limit
 * USL – Upper Specification Limit.

### 3. Useful Links
 * Six Sigma with R: <https://www.springer.com/gp/book/9781461436515>
 * Six Sigma Institute: <https://www.sixsigma-institute.org>
 * Wiki: <https://en.wikipedia.org/wiki/Six_Sigma>
 * ISO 13053-1:2011: Quantitative methods in process improvement — Six Sigma <https://www.iso.org/standard/52901.html>
 * <http://www.r-qualitytools.org>
 
## 4. Introduction
Six Sigma is one of the most widely used set of techniques, tools and method's aimed at 
improving business processes and optimizing operations.
Sigma is a statistical term which defines the standard deviation from from the mean of the normal distribution.
If we take 3 sigmas to left and 3 sigmas to the right from the mean (i.e. 6 sigmas in absolute value) then 99.99% of observations will be in this range.
Six Sigma is an indispensable tool for quality control and improvement initiatives.
It uses statistical methods to statistical methods to monitor and control a process.
And this is a strong point about Six Sigma.
By introducing verifiable data and statistical methods we can define, measure, analyse, improve, and control the business process. Six sigma involves systematic and data-driven approach to managing quality improvement projects.
The quality control management base their decisions on facts rather than assumptions and guesswork.
Indeed, Six Sigma clear focus on achieving measurable and quantifiable financial returns.
To improve existing business processes Six Sigma follows DMAIC and DMADV methodologies.
 * DMAIC is used for projects aimed at improving an existing business process.
 * DMADV is used for projects aimed at creating new product or process designs.
In this manual I will consider DMAIC cycle. DMAIC stands for:
 1.Define
 2.Measure
 3.Analyse
 4.Improve
 5.Control
In this manual I will only cover the topics of how R can be used in Measure, Analyse and Control phases.

### 5. Load the libraries
Let's first load the libraries.
```{r loading packages, message=FALSE, warning=FALSE}
library(SixSigma)
library(qcc)
library(qicharts)
library(qualityTools)
```

### 6. Loading and checking the data

Load and check the datasets.
Dataset1: pistonrings from qcc package
 * 
 * 
 
```{r load the data and pre-process them}
# Loading data from Rds file


# Checking if there are any NAs:


```

### 7. Define
During the definition phase we identify the problem area faced by the customer.
We identify customer requirements and set the project goals to achive. During this phase the PM prepares all the project initiation documents: problem statement, customer requiremets, project statement, business case, project charter, etc.
According to Six Sigma institute the define phase consists of the following 
SixSigma package offers the process mapping tool for define stage. Process map (<https://en.wikipedia.org/wiki/Business_process_mapping>)
Process map depicts the steps of the processes starting from a high, conceptual level and 
decomposing to lower levels. Each process consumes inputs X and produces output Y.
SixSigma package has a function ss.pMap() to draw the process map. However, the process of creating a map is quite tedious. I think it is much easier to create such a map in a graphical tool.
Here is an example of a process map from SixSigma package manual: 
```{r message=FALSE, warning=FALSE}
inputs.overall<-c("operators", "tools", "raw material", "facilities")
outputs.overall<-c("helicopter")
steps<-c("INSPECTION", "ASSEMBLY", "TEST", "LABELING")
#Inputs of process "i" are inputs of process "i+1"
input.output<-vector(mode="list",length=length(steps))
input.output[1]<-list(c("sheets", "..."))
input.output[2]<-list(c("sheets"))
input.output[3]<-list(c("helicopter"))
input.output[4]<-list(c("helicopter"))
#Parameters of each process
x.parameters<-vector(mode="list",length=length(steps))
x.parameters[1]<-list(c(list(c("width", "NC")),list(c("operator", "C")),
list(c("Measure pattern", "P")), list(c("discard", "P"))))
x.parameters[2]<-list(c(list(c("operator", "C")),list(c("cut", "P")),
list(c("fix", "P")), list(c("rotor.width", "C")),list(c("rotor.length",
"C")), list(c("paperclip", "C")), list(c("tape", "C"))))
x.parameters[3]<-list(c(list(c("operator", "C")),list(c("throw", "P")),
list(c("discard", "P")), list(c("environment", "N"))))
x.parameters[4]<-list(c(list(c("operator", "C")),list(c("label", "P"))))
x.parameters
#Features of each process
y.features<-vector(mode="list",length=length(steps))
y.features[1]<-list(c(list(c("ok", "Cr"))))
y.features[2]<-list(c(list(c("weight", "Cr"))))
y.features[3]<-list(c(list(c("time", "Cr"))))
y.features[4]<-list(c(list(c("label", "Cr"))))
y.features
ss.pMap(steps, inputs.overall, outputs.overall,
input.output, x.parameters, y.features,
sub="Paper Helicopter Project")

```


### 8. Measure
If you can not measure something how can you control it? 
Take the current process ("as is") and collect the relevant qualitative and quantative data about it.
Measuring phase can be divided into several steps:
### 8.1 Data types
Define what types of data to measure: qualitative or quantitative.
Qualitative data is subjective in nature and cannot be measured objectively. 
They can be nominal, ordinal (ranked or ordered) and binary.
Quantitative data is objective in nature and can be measured. They can be discrete or continuous
### 8.2 Type of statistics
Define what kind of statistics to gather. For quantitative data: 
    * measure of central tendency - mean, median, mode
    * measure of spread/dispertion: standard deviation, range

### 8.3 Data collection strategy.
    * Probability Sampling: 
      Simple Random Sampling
      Stratified Random Sampling
      Systematic Sampling
      Cluster Sampling
    * Non Probability Sampling: Convenience Sampling, Judgment Sampling, etc
    
### 8.4 Measurement System
The objective of the measure phase is to identify and understand the variation due to the
measurement system itself and variation arising from the actual production process.
Measurement System Analysis (MSA) identifies what variations come from appraiser (the operator – individual who operates the tool/instrument) and Variation from Gage due to the instrument itself. 
Variation from gage can be further classified into Accuracy, Stability, Repeatability, Reproducibility and Linearity.
The Actual Process Variation are caused by controllable factors and/or uncontrollable factors.
In general, we can calculate the observed variation as follows:
Variation (observed) = Variation (Actual) + Variation (Measurement)
Measurement system errors can be due to:
 * Accuracy – The difference between the average of observed values and the standard
 * Repeatability – Variation in measurement when a person measures the same unit repeatedly with the same measuring gage (or tool)
 * Reproducibility - Variation in measurement when two or more persons measure the same unit using the same measuring gage (or tool)
 * Stability - Variation in measurement when the same person measures the same unit using the same measuring gage (or tool) over an extended period of time.
 * Linearity – The consistency of the measurement across the entire range of the measuring gage.
Process Variation is caused by two reasons: Common Causes(noise) and Special Causes (signals).
Variation is inherent in the system and results in a stable "in control" state.
Whereas signal is an unexpected occurrenc due to unforeseen circumstances. It results in "out of control" state because the variation is not predictable.

### 8.5 Data Interpretation

Collected data can be interpreted using the following factors:
 Stability
 Normality 
 Shape
 Spread 
 Centering

Tools in R for data interpretation:
 1. Gage Repeatability and Reproducibility (Gage R&R)
 2. Run chart plots: time ordered plots of process data to understand data stability
<https://cran.r-project.org/web/packages/qicharts/vignettes/runcharts.html>  3. Standard R functionality for normality check like qqplot, statistical tests, histograms, standard deviation, mean, median, etc. 
 

### 9. Analyse
During this phase we analyse the collected data and try to find out the causes of issues.
There must be some exploitative data analysis carried out to investigate and confirm what is the causing the issues. It's an iterative approach of asking and answering questions. 

R has the following tools to perform such analysis.
 1. Cause and effect diagram
 2. Pareto chart
 3. Bar chart
 4. Histogram
 5. Box plot
 6. Scatter plot

### 11. Control

Control the process.
